---
title: "Lipid-SIP Analysis"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## How does this work?

This script is for calculation of lipid-SIP parameters that allow for quantification of biomass growth rate/turnover/generation time etc. We pair metadata with the outputs of our IRMS calibration.

# Setup

Load libraries and sourced functions

```{r}
rm(list=ls()) # Clear the environment
library(tidyverse) # CRAN v2.0.0
library(isotopia)
source("source/calculate_turnover.R") # calculate growth rates
source("source/d2H_to_F2H.R") # convert delta notation to F2H
source("source/error_prop.R") # propagate error in growth rates
source("source/rename_FAs.R") # rename FAs based on putative IDs


ref_ratio <- get_standard("2H")
ref_frac <- to_abundance(ref_ratio)
R2H_VSMOW <- ref_ratio |> as.numeric()
F2H_VSMOW <- ref_frac |> as.numeric()
```

Read in various sources of data

```{r}
# read in label dD data
label_dD <- read_rds("cache/label_dD.rds") |> select(-dD_SMOW, -at2H_label)

# read in JCR data 
JCR_metadata <- readxl::read_excel("data/JCR-metadata-Feb-2023.xlsx")|> 
  rename(F2H_label_JCR = `Effective F2H`)

# read in calibrated permafrost peak table
irms_calibrated <- readxl::read_excel("data_output/samples_gcirms_data.xlsx") |> 
  mutate(
    # parse which fraction the sample is from
    # (need to do this before cleaning sample id!)
    fraction = case_when(
      str_detect(sample_id, "F2") ~ "F2", # F2 = glycolipid
      str_detect(sample_id, "F3") ~ "F3"  # F3 = phospholipid
    ),
    # clean up the sample_id column
    sample_id = 
      case_when(
        str_detect(sample_id, "HDK") ~ sample_id,
        !str_detect(sample_id, "HDK") ~ 
                      str_extract(sample_id, ".*d-") |> 
                      str_sub(start = 1, end = -2)
      )
  ) |> 
  # parse metadata based on sample_id string
  mutate(
    site = case_when(
      str_detect(sample_id, "p35") ~ "35m",
      str_detect(sample_id, "p54") ~ "54m",
      str_detect(sample_id, "p83") ~ "83m",
      str_detect(sample_id, "pUS") ~ "Surface"
    ),
    inc_temp_C = case_when(
      str_detect(sample_id, "(-4C)") ~ -4,
      str_detect(sample_id, "(4C)") ~ 4,
      str_detect(sample_id, "(12C)") ~ 12
    ),
    experiment = case_when(
      str_detect(sample_id, "HDK23") ~ "JCR",
      TRUE ~ "Fox, AK"
    ),
    JCR_id = case_when(
      experiment == "JCR" ~ str_extract(sample_id, "HDK23-(\\d{3})"),
      TRUE ~ NA
    ),
      inc_time_d = case_when(
        str_detect(sample_id, "30d") ~ 30,
        str_detect(sample_id, "180d") ~ 180,
        str_detect(sample_id, "0d") ~ 0,
        str_detect(sample_id, "7d") ~ 7,
        experiment == "JCR" ~ 7.07 # all JCR samples ran for 7 days and change
    ),
    # parse number of C in the FAME
    n_C = abs(parse_number(compound))
  ) |> 
  select(sample_id, site, inc_temp_C, experiment, JCR_id, everything()) |> 
  # This script focuses on Fox Permafrost tunnel.
  filter(experiment == "Fox, AK")
```

Inspect the data we have

```{r}
irms_calibrated |> 
  select(sample_id, site, inc_time_d, inc_temp_C, fraction) |> 
  arrange(fraction, site, inc_temp_C, inc_time_d) |> 
  unique() |> 
  gt::gt() |> 
  gt::data_color(columns = inc_temp_C, palette = "Reds") |> 
  gt::data_color(columns = inc_time_d, palette = "Blues") |> 
  gt::data_color(columns = fraction, palette = "Greens") |> 
  gt::data_color(columns = site, palette = "Set1")
```

Load GC-FID data (see `01_GC-FID_Prep.qmd` for generation/analysis).

```{r}
FID_data <- read_rds("cache/FID_absolute_quant.RDS") |> 
  mutate(site = case_when(core == "US" ~ "Surface", TRUE ~ core)) |> 
  rename(inc_temp_C = temp, compound = `Peak Name`) |> 
  select(sample_id, compound, fraction, site, inc_temp_C, inc_time_d, abs_mass_analyte_total_ng_per_g)
```

# Inspect the data

```{r}
irms_calibrated |> 
  filter(fraction == "F3") |> 
  filter(area2 >= 2) |> 
  filter(inc_time_d < 180) |> 
  ggplot() +
  aes(
    x = inc_time_d,
    y = cal_d2H,
    color = compound
  ) +
  geom_point() +
  geom_line() +
  scale_color_viridis_d() +
  facet_wrap(vars(site, inc_temp_C), scales = "free_y") +
  labs(
    title = "Area cutoff = 2Vs, Fraction = Phospho (F3)"
  ) +
  theme_bw() +
  theme()
```

# Join with abundance data and filter

Let's generate abundance-weighted means of the data to see how the community d2H increases over time. We can then use these to calculate assemblage-level biomass turnover rates.

We'll apply an area filter of \> 2Vs.

```{r}
irms_w_FID <- irms_calibrated |> 
  left_join(FID_data, by = c("sample_id", "site", "inc_temp_C", "inc_time_d", "compound", "fraction")) |> 
  # pare down the data
  select(-c(folder, rep, rt_start, rt_end, amp2, amp3, area3, bgrd2_end, 
            bgrd3_start,  bgrd3_end, `r3/2`, `rd3/2`, `d3/2`, n_overlapping, 
            n_matches, ref_ratio_at_rt, rel_area, mean_amp_identified,
            ref_peak_outliers)) |> 
  # apply area filter
  filter(area2 >= 2)
```

Inspect the data

```{r}
irms_w_FID |>
  ggplot() +
  aes(
    x = cal_d2H,
    y = abs_mass_analyte_total_ng_per_g, label = paste(sample_id, compound),
    color = as.factor(inc_time_d),
    shape = site
  ) +
  geom_point() +
  theme_bw()
```

# Visualize differences between memory corrected data and non-corrected data

```{r}
irms_cal_all <- readxl::read_excel("data_output/samples_gcirms_data_all.xlsx") |> 
  mutate(
    # parse which fraction the sample is from
    # (need to do this before cleaning sample id!)
    fraction = case_when(
      str_detect(sample_id, "F2") ~ "F2", # F2 = glycolipid
      str_detect(sample_id, "F3") ~ "F3"  # F3 = phospholipid
    ),
    # clean up the sample_id column
    sample_id = 
      case_when(
        str_detect(sample_id, "HDK") ~ sample_id,
        !str_detect(sample_id, "HDK") ~ 
                      str_extract(sample_id, ".*d-") |> 
                      str_sub(start = 1, end = -2)
      )
  ) |> 
  # parse metadata based on sample_id string
  mutate(
    site = case_when(
      str_detect(sample_id, "p35") ~ "35m",
      str_detect(sample_id, "p54") ~ "54m",
      str_detect(sample_id, "p83") ~ "83m",
      str_detect(sample_id, "pUS") ~ "Surface"
    ),
    inc_temp_C = case_when(
      str_detect(sample_id, "(-4C)") ~ -4,
      str_detect(sample_id, "(4C)") ~ 4,
      str_detect(sample_id, "(12C)") ~ 12
    ),
    experiment = case_when(
      str_detect(sample_id, "HDK23") ~ "JCR",
      TRUE ~ "Fox, AK"
    ),
    JCR_id = case_when(
      experiment == "JCR" ~ str_extract(sample_id, "HDK23-(\\d{3})"),
      TRUE ~ NA
    ),
      inc_time_d = case_when(
        str_detect(sample_id, "30d") ~ 30,
        str_detect(sample_id, "180d") ~ 180,
        str_detect(sample_id, "0d") ~ 0,
        str_detect(sample_id, "7d") ~ 7,
        experiment == "JCR" ~ 7.07 # all JCR samples ran for 7 days and change
    ),
    # parse number of C in the FAME
    n_C = abs(parse_number(compound))
  ) |> 
  select(sample_id, site, inc_temp_C, experiment, JCR_id, everything()) |> 
  # This script focuses on Fox Permafrost tunnel.
  filter(experiment == "Fox, AK")

irms_cal_all |> 
  filter(area2 >= 2 & area2 < 20) |> 
  filter(!is.na(cal_d2H_yes_mem)) |> 
  mutate(diff_mem_nomem = cal_d2H_no_mem - cal_d2H_yes_mem) |> 
  ggplot(
    aes(
      y = diff_mem_nomem,
      x = area2,
      color = as.factor(inc_time_d)
    )
  ) +
  geom_point() +
  labs(
    y = "d2H no correction - d2H memory corrected"
  ) +
  theme_bw()
```

# Calculate abundance-weighted mean

We can aggregate our growth rate measurements into assemblage-level estimates of microbial community growth by calculating an abundance-weighted mean $(\bar{x})$:

$$
\bar{x} = \frac{ \sum_{i = 1}^{N} x_iw_i}{ \sum_{i = 1}^{N} w_i }
$$

Where $x_i$ is the isotopic composition of a compound $i$ and $w_i$ is the relative abundance (weighting) of compound $i$. Here, our weight (w) is the compound abundance determined by GC-FID: `abs_mass_analyte_total_ng_per_g` and our data (x) is the calibrated d2H.

```{r}
irms_w_FID_wm <- irms_w_FID |> 
  group_by(fraction, site, inc_temp_C, inc_time_d) |> 
  dplyr::summarise(
    cal_d2H_wm = 
      sum(cal_d2H * abs_mass_analyte_total_ng_per_g, na.rm = TRUE) /
      sum(abs_mass_analyte_total_ng_per_g, na.rm = TRUE),
    .groups = "keep"
  )


# visualize the weighted mean data
irms_w_FID_wm |> 
  filter(inc_time_d < 180) |> 
  filter(fraction != "F2") |> 
  filter(site != "Surface") |> 
  ggplot() +
  aes(
    x = inc_time_d,
    y = cal_d2H_wm
  ) +
  geom_point() +
  geom_line() +
  facet_wrap(vars(site, inc_temp_C))
```

# Calculate growth rate

## Calculate mean F2H of t0 compounds

```{r}
sample_peak_table_t0 <- sample_peak_table |> 
  filter(inc_time_d == 0) |> 
  group_by(compound, site) |> 
  dplyr::summarize(mean_F2H_t0 = mean(F2H, na.rm = TRUE)) |> 
  ungroup() |> 
  filter(!is.na(compound))

sample_FA_t0_mean <- sample_peak_table_t0 |> 
  pull(mean_F2H_t0) |> mean()
```

### Adjust label strengths

```{r}
sample_peak_table_FL <- sample_peak_table |> 
  left_join(label_dD, by = "sample_id") |> 
  mutate(F2H_label = case_when(
    is.na(F2H_label) ~ F2H_label_JCR,
    TRUE ~ F2H_label
    )
  ) |> select(-F2H_label_JCR)
```

### Growth Calcs

```{r}
growth_data <- sample_peak_table_FL |> 
  #add in the t0 data
  left_join(sample_peak_table_t0, by = c("compound", "site")) |> 
  mutate(
    mean_F2H_t0 = case_when(
      experiment == "JCR" ~ 0.00014247086, # mean natabund from Caro et al. 2023
      experiment == "Fox, AK" ~ mean_F2H_t0
    ),
    # for peaks not ID'd in t0 sample, use mean value of all t0 FAs
    mean_F2H_t0 = case_when(
      is.na(mean_F2H_t0) ~ sample_FA_t0_mean,
      TRUE ~ mean_F2H_t0
    )
  ) |> 
  mutate(
    u_d = case_when(
      inc_time_d == 0 ~ 0, # growth rate of organisms at t0 is zero
      inc_time_d > 0 ~
        calculate_turnover(
        a = 0.71,
        FT = F2H,
        F0 = mean_F2H_t0,
        FL = F2H_label,
        t = inc_time_d
        )
    )
  ) |> 
  rename_FAs() # update FA names based on putative identification
```

Inspect the data

```{r}
growth_data |> 
  select(sample_id, compound, u_d, F2H, mean_F2H_t0, F2H_label) |> 
  filter(is.na(u_d)) |> 
  gt::gt()
```

# Join with abundance data

First let's make a smaller/tidier version of growth_data that is more amenable to joining:

```{r}
growth_data_tidy <- growth_data |> 
  filter(experiment == "Fox, AK") |> # ignore JCR samples
  select(
    sample_id,
    site,
    inc_temp_C,
    inc_time_d,
    fraction,
    compound, 
    n_C,
    d2H, at2H, F2H,
    u_d
  )

growth_w_FID <- growth_data_tidy |> 
  left_join(FID_data, by = c("sample_id", "site", "inc_temp_C", "inc_time_d", "compound", "fraction"))
```

## Abundance-weighted mean

We can aggregate our growth rate measurements into assemblage-level estimates of microbial community growth by calculating an abundance-weighted mean $(\bar{x})$:

$$
\bar{x} = \frac{ \sum_{i = 1}^{N} x_iw_i}{ \sum_{i = 1}^{N} w_i }
$$

Where $x_i$ is the isotopic fractional abundance of a compound $i$ and $w_i$ is the relative abundance (weighting) of compound $i$. Here, our weight (w) is the compound abundance determined by GC-FID: `mass_analyte_ng_per_g.FID` and our data (x) is the growth rate `u_d`.

```{r}
weighted_mean_growth_data <- growth_w_FID |> 
  dplyr::group_by(site, inc_temp_C, inc_time_d, fraction) |> 
  dplyr::summarise(
    u_d = sum(u_d * mass_analyte_ng_per_g.FID, na.rm = TRUE) /
      sum(mass_analyte_ng_per_g.FID, na.rm = TRUE),
    .groups = "keep"
  ) |> 
  ungroup() |> 
  mutate(
    gen_d = log(2) / u_d, # calculate weighted mean gen time
    compound = "Weighted Mean", # plotting identifier
    )
```

# Plots

### Lipid abundance

```{r}
PLFA_sums <- growth_w_FID |> 
  filter(fraction == "phospho-IPL") |>
  group_by(site, inc_time_d, inc_temp_C) |> 
  summarize(
    PLFA_total_ng_per_g = sum(mass_analyte_ng_per_g.FID, na.rm = TRUE)
  )

PLFA_sums |> 
  ggplot(
    aes(
      x = inc_time_d,
      y = PLFA_total_ng_per_g,
      color = as.factor(inc_temp_C)
    )
  ) +
  geom_point() +
  geom_line() +
  scale_color_viridis_d() +
  facet_wrap(vars(site), scales = "free_y")
```

## Fox prelim

### Compound Specific µ with Mean

```{r}
p_ud_weighted <- growth_data |> 
  filter(experiment == "Fox, AK") |> # filter out JCR
  filter(!is.na(site)) |> # filter out site NA
  # Add in weighted mean data:
  bind_rows(weighted_mean_growth_data) |>
  filter(site != "Surface") |> # plot surface sample separately
  filter(inc_time_d == 30) |> # choose incubation time to focus on
  filter(fraction == "phospho-IPL") |> # choose fraction
  # factors helpful for plotting:
  mutate(
    inc_temp_C_str = factor(paste0(inc_temp_C, "C"), c("-4C", "4C", "12C")),
    is_mean = if_else(compound == "Weighted Mean", true = TRUE, false = FALSE),
    compound = str_replace(compound, "ω", "w")
    ) |> 
  # Plot it:
  ggplot(
    aes(
      # reorder y variables in ascending order within each facet
      y = tidytext::reorder_within(
        compound, u_d, list(site, inc_temp_C)
        ),
      x = u_d
    )
  ) +
  geom_point(
    aes(shape = inc_temp_C_str, fill = inc_temp_C_str),
    stroke = 1, size = 2,
    color = "black"
    ) +
  scale_shape_manual(values = c(21, 22, 23)) +
  ggrepel::geom_text_repel(
    data = function(df) df |> filter(compound == "Weighted Mean"),
    aes(
      label = paste(formatC(u_d, format = "e", digits = 1), 
                    " = \n", round(gen_d, 0), "days")
      ),
    nudge_x = 0.001,
    nudge_y = -1,
    color = "red",
    size = 3,
    segment.alpha = 0.5, segment.curvature = -1e-20
    
  ) +
  scale_fill_manual(values = c("#3e0751", "#468e8c", "#f9e955")) +
  facet_wrap(vars(inc_temp_C_str, site), ncol = 3, scales = "free_y") +
  tidytext::scale_y_reordered() +
  labs(
    y = "", x = latex2exp::TeX("Compound-Specific Growth Rate $(d ^{-1})$")
  ) +
  guides(
    color = "none",
    shape = "none",
    fill = "none"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dotted", color = "gray"),
    strip.background = element_rect(fill = "black", color = "black"),
    strip.text = element_text(color = "white", face = "bold")
  )
p_ud_weighted 

# cowplot::save_plot(
#   filename = "fig_output/p_ud_weighted.pdf",
#   plot = p_ud_weighted,
#   base_height = 9,
#   base_width = 12
# )
```

Make the same plot but with the surface samples:

```{r}
p_ud_weighted_surface <- growth_data |> 
  filter(experiment == "Fox, AK") |> # filter out JCR
  filter(!is.na(site)) |> # filter out site NA
  # Add in weighted mean data:
  bind_rows(weighted_mean_growth_data) |>
  filter(site == "Surface") |> # plot surface sample separately
  filter(inc_time_d == 30) |> # choose incubation time to focus on
  filter(fraction == "phospho-IPL") |> # choose fraction
  # factors helpful for plotting:
  mutate(
    inc_temp_C_str = factor(paste0(inc_temp_C, "C"), c("-4C", "4C", "12C")),
    is_mean = if_else(compound == "Weighted Mean", true = TRUE, false = FALSE),
    compound = str_replace(compound, "ω", "w")
    ) |> 
  # Plot it:
  ggplot(
    aes(
      # reorder y variables in ascending order within each facet
      y = tidytext::reorder_within(
        compound, u_d, list(site, inc_temp_C)
        ),
      x = u_d
    )
  ) +
  geom_point(
    stroke = 1, size = 2, fill = "black",
    color = "black"
    ) +
  ggrepel::geom_text_repel(
    data = function(df) df |> filter(compound == "Weighted Mean"),
    aes(
      label = paste(formatC(u_d, format = "e", digits = 1), 
                    " = \n", round(gen_d, 0), "days")
      ),
    nudge_x = 0.01,
    nudge_y = -1,
    color = "red",
    size = 3,
    segment.alpha = 0.5, segment.curvature = -1e-20
    
  ) +
  #facet_wrap(vars(inc_temp_C_str, site), ncol = 3, scales = "free_y") +
  tidytext::scale_y_reordered() +
  #coord_cartesian(xlim = c(0, 0.25)) +
  labs(
    y = "", x = latex2exp::TeX("Compound-Specific Growth Rate $(day ^{-1})$"),
    title = "Surface Permafrost"
  ) +
  guides(
    color = "none",
    shape = "none",
    fill = "none"
  ) +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), 
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dotted", color = "gray"),
    strip.background = element_rect(fill = "white", color = "white"),
    aspect.ratio = 1,
    plot.title = element_text(hjust = 0.5)
  )
p_ud_weighted_surface 

# cowplot::save_plot(
# filename = "fig_output/p_ud_weighted_surface2.pdf",
# plot = p_ud_weighted_surface,
# base_height = 5,
# base_width = 4
# )
```

#### F2H

```{r}
p_fox_F2H <- growth_data |> 
  filter(experiment == "Fox, AK") |> 
  filter(site != "Surface") |> 
  filter(!is.na(site)) |> 
  filter(inc_time_d != 180) |> 
  ggplot() +
  aes(
    #y = F2H * 1e6, # uncomment this for F2H in ppm
    y = as.numeric(d2H), # using delta notation for now
    x = inc_time_d,
    color = compound,
    shape = fraction
  ) +
  geom_point() +
  facet_grid(site~as.factor(inc_temp_C)) +
  labs(
    x = "Incubation time (days)",
    y = latex2exp::TeX("$\\delta^{2}H_{VSMOW}$")
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(color = "white", face = "bold"),
    strip.background = element_rect(color = "black", fill = "black")
  )
p_fox_F2H

# anomalously high d2H is probably an area effect. future data calibration/scrubbing will most likely remove these!

# cowplot::save_plot(
#   filename = "fig_output/p_fox_F2H.pdf",
#   plot = p_fox_F2H,
#   base_height = 10,
#   base_width = 10
# )
```

```{r}
p_fox_F2H_surface <- growth_data |> 
  filter(experiment == "Fox, AK") |> 
  filter(site == "Surface") |> 
  filter(!is.na(site)) |>
  mutate(
    diff_F2H = case_when(
      inc_time_d == 0 ~ 0,
      inc_time_d > 0 ~ F2H - mean_F2H_t0
    )
  ) |> 
  ggplot() +
  aes(
    y = diff_F2H * 1e6,
    x = inc_time_d,
    shape = fraction,
    group = compound
  ) +
  geom_point() +
  geom_line() +
  facet_grid(vars(fraction)) +
  labs(
    x = "Incubation Time (Days)",
    y = latex2exp::TeX("$\\Delta^{2}F_{biomass}$  (ppm)"),
    title = "Surface permafrost (10 cm depth)"
  ) +
  theme_bw()
p_fox_F2H_surface

# cowplot::save_plot(
#   filename = "fig_output/p_fox_F2H_surface.pdf",
#   plot = p_fox_F2H_surface,
#   base_height = 4,
#   base_width = 4
# ) 
```

#### Growth rate (µ)

```{r}
p_fox_u_d_surface <- growth_data |> 
  filter(experiment == "Fox, AK") |> 
  filter(site == "Surface") |> 
  filter(!is.na(site)) |> 
  filter(inc_time_d !=0) |> 
  filter(!is.na(u_d)) |> 
  mutate(inc_time_d_str = factor(paste(inc_time_d, "days"), 
                                 levels = c("7 days", "30 days")),
         compound = forcats::fct_reorder(compound, n_C, .desc = TRUE)) |> 
  ggplot() +
  aes(
    y = compound,
    x = u_d
  ) +
  geom_segment(
    aes(yend = compound),
    xend = 0,
  ) +
  geom_point(aes(color = fraction, shape = fraction),
             size = 3) +
  scale_color_brewer(palette = "Set1")+
  facet_wrap(vars(inc_time_d_str, fraction), scales = "free_x") +
  coord_cartesian(expand = FALSE, ylim = c(0, 17), xlim = c(0, 0.023)) +
  scale_x_continuous(
    sec.axis = sec_axis(
      trans = ~ log(2) / .,
      breaks = c(1000, 360, 180, 120, 80, 60, 35),
      name = "Generation Time (d)")
  ) +
  labs(
    x = latex2exp::TeX("Biomass Growth Rate $(d ^{-1})$"),
    y = "",
    title = "Surface permafrost (10 cm depth)",
    color = "Incubation Time",
    shape = "Incubation Time"
  ) +
  ggprism::annotation_ticks() +
  theme_classic() +
  theme(
    strip.background = element_rect(fill = "black", color = "black"),
    strip.text = element_text(color = "white", face = "bold"),
    panel.border = element_rect(color = "black", fill = NA),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted"),
    axis.text.x.top = element_text(angle = 45, hjust = 0),
    axis.text.x.bottom = element_text(angle = 45, hjust = 1),
    strip.placement = "outside",
    strip.background.x = element_blank(),
    strip.text.x = element_blank(),
    aspect.ratio = 1,
    axis.text = element_text(color = "black")
  )
p_fox_u_d_surface

# cowplot::save_plot(
#   filename = "fig_output/p_fox_u_d_surface.pdf",
#   plot = p_fox_u_d_surface,
#   base_height = 10,
#   base_width = 10
# )
```

```{r}
p_fox_subsurface_u_d <- growth_data |> 
  filter(experiment == "Fox, AK") |> 
  filter(site != "Surface") |> 
  filter(!is.na(site)) |> 
  filter(inc_time_d == 7) |> # only look at 30 day incubations!
  filter(!is.na(u_d)) |> 
  mutate(inc_time_d_str = factor(paste(inc_time_d, "days"), 
                                 levels = c("7 days", "30 days")),
         compound = forcats::fct_reorder(compound, n_C, .desc = TRUE)) |> 
  ggplot() +
  aes(
    x = u_d,
    y = compound,
    color = as.factor(inc_temp_C),
    shape = fraction
  ) +
  geom_segment(
    aes(yend = compound),
    xend = 0,
    color = "black"
  ) +
  geom_point() +
  facet_grid(site~inc_temp_C) +
  scale_color_viridis_d(option = "plasma", end = 0.7) +
  scale_x_continuous(
    sec.axis = sec_axis(
      trans = ~ log(2) / .,
      breaks = c(3650, 1200, 800, 600, 500, 365, 120),
      name = "Generation Time (d)")
  ) +
  #coord_cartesian(xlim = c(0, 0.0008), ylim = c(0,13)) +
  ggprism::annotation_ticks() +
  labs(
    x = latex2exp::TeX("Biomass Growth Rate $(d ^{-1})$"),
    y = "", 
    color = "Incubation Temperature (˚C)", shape = "Incubation Temperature (˚C)"
  ) +
  theme_classic() +
  theme(
    strip.background = element_rect(fill = "black", color = "black"),
    strip.text = element_text(color = "white", face = "bold"),
    panel.border = element_rect(color = "black", fill = NA),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted"),
    axis.text.x.top = element_text(angle = 45, hjust = 0),
    axis.text.x.bottom = element_text(angle = 45, hjust = 1),
    strip.placement = "outside",
    strip.background.x = element_blank(),
    strip.text.x = element_blank(),
    aspect.ratio = 1,
    axis.text = element_text(color = "black")
  )
p_fox_subsurface_u_d

# cowplot::save_plot(
#   filename = "fig_output/p_fox_subsurface_u_d.pdf",
#   plot = p_fox_subsurface_u_d,
#   base_width = 10,
#   base_height = 10
# )
```

### Growth vs. abundance

Does lipid abundance correlate with its production rate?

```{r}
p_u_d_vs_abund <- growth_w_FID |> 
  filter(site != "Surface") |> 
  mutate(inc_temp_C_str = factor(paste(inc_temp_C, "˚C"), c("-4 ˚C", "4 ˚C", "12 ˚C"))) |> 
  mutate(inc_time_d_str = factor(paste(inc_time_d, "days"), c("7 days", "30 days", "180 days"))) |> 
  filter(inc_time_d != 0) |> 
  ggplot(
    aes(
      x = mass_analyte_ng_per_g.FID,
      y = u_d,
      shape = site,
      color = site,
      label = compound
    )
  ) +
  geom_point(alpha = 1) +
  scale_color_brewer(palette = "Set1") +
  facet_grid(inc_time_d_str~inc_temp_C_str, scales = "free_y") +
  labs(
    x = "Lipid abundance (ng/g permafrost)",
    y = latex2exp::TeX("Growth rate $(day^{-1})$"),
    color = "Site",
    shape = "Site"
  ) +
  ggprism::annotation_ticks(sides = "bl") +
  theme_bw() +
  theme(
    axis.text = element_text(color = "black", face = "bold"),
    strip.text = element_text(color = "white", face = "bold", size = 16),
    strip.background = element_rect(color = "black", fill = "black"),
    panel.grid = element_blank(),
    aspect.ratio = 1,
    axis.ticks = element_blank()
  )
p_u_d_vs_abund
# 
# cowplot::save_plot(
#   filename = "fig_output/p_u_d_vs_abund.pdf",
#   plot = p_u_d_vs_abund,
#   base_height = 10,
#   base_width = 10
# )

# ggsave(
#   plot = p_u_d_vs_abund,
#   filename = "fig_output/p_u_d_vs_abund.png",
#   height = 10,
#   width = 10
# )
```

```{r}
p_surface_u_d_abund <- growth_w_FID |> 
  filter(site == "Surface") |> 
  mutate(inc_temp_C_str = factor(paste(inc_temp_C, "˚C"), c("-4 ˚C", "4 ˚C", "12 ˚C"))) |> 
  mutate(inc_time_d_str = factor(paste(inc_time_d, "days"), c("7 days", "30 days", "180 days"))) |> 
  ggplot(
    aes(
      x = mass_analyte_ng_per_g.FID,
      y = u_d,
      label = compound,
      shape = inc_time_d_str,
      color = inc_time_d_str
    )
  ) +
  geom_point(alpha = 1) +
  labs(
    x = "Lipid abundance (ng/g permafrost)",
    y = latex2exp::TeX("Growth rate $(day^{-1})$"),
    color = "Site",
    shape = "Site"
  ) +
  ggprism::annotation_ticks(sides = "bl") +
  theme_bw() +
  theme(
    axis.text = element_text(color = "black", face = "bold"),
    strip.text = element_text(color = "white", face = "bold", size = 16),
    strip.background = element_rect(color = "black", fill = "black"),
    panel.grid = element_blank(),
    aspect.ratio = 1,
    axis.ticks = element_blank()
  )
p_surface_u_d_abund

# cowplot::save_plot(
#   filename = "fig_output/p_surface_u_d_abund.pdf",
#   plot = p_surface_u_d_abund,
#   base_height = 4,
#   base_width = 4
# )
```

## JCR prelim

```{r}
p_JCR_u_d <- growth_data |> 
  filter(experiment == "JCR") |> 
  mutate(compound = forcats::fct_reorder(compound, n_C, .desc = TRUE)) |> 
  ggplot() +
  aes(
    x = u_d,
    y = compound,
    fill = Site,
  ) +
  geom_segment(
    aes(
      yend = compound,
      xend = u_d
    ),
    xend = 0,
    color = "black"
  ) +
  geom_point(
    shape = 21,
    color = "black",
    size = 3
  ) +
  scale_fill_brewer(palette = "Set1") +
  scale_x_continuous(
    sec.axis = sec_axis(
      trans = ~ log(2) / .,
      breaks = c(10, 18, 30, 90, 1000),
      name = "Generation Time (d)")
  ) +
  #coord_cartesian(expand = FALSE, xlim = c(0, 0.07), ylim = c(0, 17)) +
  facet_wrap(vars(`Site`), ncol = 2) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(), 
    panel.grid.major.y = element_line(linetype = "dotted", color = "gray"),
    strip.background = element_rect(fill = "#8a1500", color = NA),
    strip.text = element_text(color = "white", face = "bold"),
    axis.text = element_text(color = "black"),
    axis.text.x.bottom = element_text(angle = 45, hjust = 1),
    axis.text.x.top = element_text(angle = 45, hjust = 0),
    panel.spacing = unit(0.5, "cm"),
    axis.text.y = element_text(size = 6)
  ) +
  labs(
    x = latex2exp::TeX("Biomass Growth Rate $(d ^{-1})$"),
    y = "",
    caption = latex2exp::TeX("Secondary axis represents apparent biomass generation time, where $T_G = ln(2)/\\mu_d$")
  )
p_JCR_u_d

```
